Multi-level modelling with Openflexo/FML - A contribution to the \mpc

### 2021-05-28 4-Technology.tex
Updates from Overleaf
synchro from overleaf
replace OMG by macro

To meet the \mpc, we have decided to use the general
approach of model federation~\cite{Golra2016-federation}. \emph{Model
  federation} is a way to assemble models using some kind of
low-coupling links. It has been studied to answer the OMG' RFP
Semantic Modeling for Information Federation (SIMF)~\cite{simf}. In
contrast to approaches that compose metamodels into a single large
metamodel grouping all needed entities, model federation build links
among models and metamodels (even through levels) to make ``things''
work together. As an illustration of this feature, we developed a
free-modeling editor -- freeing oneself from the bonds of
model/metamodel conformity -- that is presented
in~\cite{models2016-freemodel}. Another very interesting feature of
this approach is the strong decoupling among tools that remain usable
after federations are made.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{Figures/FMLCoreModel.pdf}
    \caption{The FML core metamodel}
    \label{fig:mm}
\end{figure}

FML simplified metamodel is provided in Figure~\ref{fig:mm}, it is designed to define virtual models. A virtual model
is composed of a set of \emph{concepts}, while itself being a concept.
Hence, virtual models are structuring units while concepts are the
core entities. A concept has a set of \emph{roles} and
\emph{behaviors}. A parallel to object-oriented approach can be useful
to understand FML\footnote{Some aspects of FML do not exist in object oriented approach.}. A concept corresponds to a class, its roles to the
attributes of the class and its behaviors to the methods of the class.
These roles have types defining the kind of value the role will point
at runtime.
Whenever a type external to the federation space is used, one needs to use a \emph{model slot}. A model
slot is a mediation entity in charge of giving
access to external elements using
a \emph{technology adapter}\footnote{It is a reusable library that defines
connections between the FML execution engine and a particular
technological space.}.

>>>>

To meet the \mpc, we have decided to use the general
approach of model federation~\cite{Golra2016-federation}. \emph{Model
  federation} is a way to assemble models using some kind of
low-coupling links. It has been studied to answer the \OMG' RFP
Semantic Modeling for Information Federation (SIMF)~\cite{simf}. In
contrast to approaches that compose metamodels into a single large
metamodel grouping all needed entities, model federation build links
among models and metamodels (even through levels) to make ``things''
work together. As an illustration of this feature, we developed a
free-modeling editor -- freeing oneself from the bonds of
model/metamodel conformity -- that is presented
in~\cite{models2016-freemodel}. Another very interesting feature of
this approach is the strong decoupling among tools that remain usable
after federations are made.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{Figures/FMLCoreModel.pdf}
    \caption{The FML core metamodel}
    \label{fig:mm}
\end{figure}
%\noteAntoine{Dans la figure ne manque-t-il pas les containements ? (VM vers Concepts)}

FML simplified metamodel is provided in Figure~\ref{fig:mm}. It is designed to reify models as virtual models. A virtual model
is composed of a set of \emph{concepts}, while itself being a concept.
Hence, virtual models are structuring units while concepts are the
core entities. A concept has a set of \emph{roles} and
\emph{behaviors}. A parallel to object-oriented approach can be useful
to understand FML\footnote{Some aspects of FML do not exist in object oriented approach.}. A concept corresponds to a class, its roles to the
attributes of the class and its behaviors to the methods of the class.
These roles have types defining the kind of value the role will point
at runtime.
Whenever a type external to the federation space is used, one needs to use a \emph{model slot}. A model
slot is a mediation entity in charge of giving
access to external elements using
a \emph{technology adapter}\footnote{It is a reusable library that defines
connections between the FML execution engine and a particular
technological space.}.
